---
title: "ZTZ method"
author: "Xiaoxia Champon"
date: "9/16/2021"
output: html_document
---
---
title: 'ST758 Homework 3'
author: "Xiaoxia Champon"
date: "9/8/2021"
output: pdf_document
---
I have neither given nor received  unauthorized aid on this test for assignment. Xiaoxia Champon
```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = TRUE)
library(tidyr)
library(ggplot2)
library(dplyr)
library(BBmisc)
library(wordspace)
library(matlib)
```
survey for the project is completed.
Question 2-5 is on paper and uploaded in another PDF document.
1.
quick sort
```{r}
set.seed(123)
quicksort <- function(x){#find pivot position for a number
  n = length(x)
  if(n<=1){
    return(x)
  }
  index  <- sample(seq(1,n,1),1)  #randomly choose a number
  
  #put the number in the beginning of the sequence
  if(index==n){
    neworder <- c(n, 1:(n-1))
    x <- x[neworder] 
  } else if(index>1){
      neworder <- c(index,1:(index-1),(index+1):n) 
      x <- x[neworder]
  }
  
  i=1
  for(j in 1:(n-1)){
    if (x[j+1] < x[1]){
      xi1 <- x[i+1]
      xj1 <- x[j+1]
      x[i+1] <- xj1
      x[j+1] <- xi1
      i = i+1   #update i by 1
    }
  }
  #swap x[1] and x[i]
  x1 <- x[1]
  xi <- x[i]
  x[1] <- xi
  x[i] <- x1
  
  if(i==1){
    low.half = x[1]
    high.half = quicksort(x[2:n])
    return(c(low.half, high.half))
  } else if (i==n){
    low.half = quicksort(x[1:(n-1)])
    high.half = x[n]
    return(c(low.half, high.half))
  } else{
    low.half = quicksort(x[1:(i-1)])
    high.half = quicksort(x[(i+1):n])
    return(c(low.half, x[i], high.half))
  }
  
   #returns list of vector x with element i in right place
  
}
```
try different n
```{r, echo=FALSE}
set.seed(123)
for(n in c(1000, 5000, 10000, 20000)){
  timex = numeric(20)
  timey = numeric(20)
  timez = numeric(20)
  
  for(i in 1:20){
    x = seq(1,n,1)
    y = seq(n,1,-1)
    z = sample(seq(1,n,1), replace=FALSE)
  
    tx = proc.time()[3]
    storage = quicksort(x)
    timex[i] = as.numeric(proc.time()[3]-tx)
    
    ty = proc.time()[3]
    storage = quicksort(y)
    timey[i] = as.numeric(proc.time()[3]-ty)
    
    tz = proc.time()[3]
    storage = quicksort(z)
    timez[i] = as.numeric(proc.time()[3]-tz)
  }
  
  print(paste("The mean sorting time for x at n=", n, "is", round(mean(timex),3)))
  print(paste("The mean sorting time for y at n=", n, "is", round(mean(timey),3)))
  print(paste("The mean sorting time for z at n=", n, "is", round(mean(timez),3)))
  print("")

  
}

```
      | $n=1000$ | $n=5000$ | $n=10000$ | $n=20000$
------|----------|----------|-----------|-----------
$x$   | 0.026    | 0.117    | 0.238     | 0.49
$y$   | 0.023    | 0.120    | 0.237     | 0.489
$z$   | 0.023    | 0.120    | 0.238     | 0.489
It takes almost the same amout of the time. The three way of sorting are the same order. \newline\newline

6 simulation:The relationship between minimum singular value and maximum VIF is not true for the original data but it's true for the unit length data.
```{r}
set.seed(123)
eigenmin=c(1)
maxvif=c(1)
eigenminu=c(1)
maxvifu=c(1)
index = 1

for (p in 3:5){

for (m in 1:1000){
  
  for (n in seq(15,200,by=5)){
    for (gam in seq(0,0.95,by=0.05)){
      
      w=matrix(rnorm(n*p,mean=10,sd=100), nrow=n, byrow=T)
      x=matrix(0,n,p)
      x[,1]=c(rep(1,n))
      unitx=matrix(0,n,p)
      unitx[,1]=c(rep(1,n))
  
      for ( i in 2:p){
      x[,i]=sqrt(1-gam^2)*w[,i]+gam*w[,p]
      
      }
      stx=scale(x[,-1],center=TRUE,scale=TRUE)
      z=stx/sqrt(n)
      VIF=max(diag(solve(t(z)%*%z)))
      
      unitx=normalize.cols(x,method ="euclidean" )
      stu=scale(unitx[,-1],center=TRUE,scale=TRUE)
      zu=stu/sqrt(n)
      VIFU=max(diag(solve(t(zu)%*%zu)))
     
      
      eigenmin[index]=min(eigen(t(x)%*%x)$values)
      maxvif[index]=VIF
  
      eigenminu[index]=min(eigen(t(unitx)%*%unitx)$values)
      maxvifu[index]=VIFU
      index = index + 1
      
    }
    }
  

}
plot(maxvif,eigenmin,xlab="Maximum VIF", ylab="Minimum eigenvalue",main=bquote("Original data"),xlim=c(0, max(maxvif)+10), ylim=c(-1,max(eigenmin)+10 ))
abline(v=10)
abline(h=0.1)

plot(maxvifu,eigenminu,xlab="Maximum VIF", ylab="Minimum eigenvalue",main="Unit Length data",xlim=c(0, max(maxvifu)+1), ylim=c(-1,max(eigenminu)+1 ))
abline(v=10)
abline(h=0.1)
  
}
```



